Understanding Parameterized Quantum Circuit Learning for Quantum Chemical Applications

### 2025-04-11 manuscript/main.tex
update QiskitRegressor.py
upload manuscript

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE5_RUDAL_lineplot}
		\caption{}
		\label{fig:bse5RUDAL_lineplot}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_RUDAL_lineplot}
		\caption{}
		\label{fig:ddccRUDAL_lineplot}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.65\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/distribution_parity}
		\caption{}
		\label{fig:BSE5_distribution_parity}
	\end{subfigure}
	\hfill

	\begin{subfigure}[b]{0.65\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/distribution_parity}
		\caption{}
		\label{fig:ddccdistribution_parity}
	\end{subfigure}	
	\caption{Line plots to highlight the performance as the number of ansatz layers (AL; left) and re-upload depth (RUD; right) increases for the 5 qubits (a) BSE49 and (b) DDCC data. Additionally, regression parity plots for each AL and RUD are shown for the (c) BSE49 and (d) DDCC data, where the x-axis denotes the predicted target values and the y-axis denotes the reference data.}
	\label{fig:BSE_distribution_parity}	
\end{figure}

@@

1 iteration: 25004.5437 s

@@

5 qubit/5 ansatz layers: 2*5 + 3 * 5 * 5 = 85 parameters (75 trainable, 5 * 2 features)
A2 $2n$ parameters
HWE-CNOT $3nL$ parameters
2 qubit gates $nL$
Number of parameters (2n+3nL)=(2+3L)*n

$n$, number of qubits and $L$, number of circuit layers

To efficiently run DDCC on IBM Quebec, splitting the data into batches of $\approx 4$ samples with 64 $t_{2}$-amplitudes each.

One iteration would require approximately $N_{\text{samples}} * N_{occ}^{2} * N_{virt}^{2} * N_{\text{shots}} * N_{\text{observables}}$ circuit executions (+ whatever SPSA costs to run per iteration)

Ran using the state vector model parameters for one iteration to test the optimization and resilience levels using Fake Quebec before running on the real device

$\{1024 \times x \vert  x \in [1,10]\}$

Fake
Optimization level 2 
resilience level 0

Regarding the number of circuit executions vs performance 3072 (1024 times 3) is the best number of shots

Fig. \ref{fig:optresfakequebec}
Fig. \ref{fig:shotsfakequebec}

Real
Optimization level 2 
resilience level 1
3072

resilience level 2 is too expensive on the real device
Regarding the number of circuit executions vs performance 3072 (1024 times 3) is the best number of shots

Error handling for real run with iter>1

For the experiments using the \textit{FakeQuebec} and \textit{ibm\_quebec} backends, we performed small tests, initialized using the state-vector parameters with one iteration, to determine the best circuit optimization and error mitigation (resilience) levels.

optimization levels
none (0), light (1), medium (2), high (3)

resilience levels (error mitigation)
none (0) 
level 1 readout error mitigation and measurement twirling using Twirled Readout Error eXtinction (TREX) \cite{van_den_berg_model-free_2022}
level 2 level 1 + gate twirling and zero noise extrapolation (ZNE)\cite{kandala_error_2019,li_efficient_2017,temme_error_2017}

>>>>

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE5_RUDAL_lineplot}
		\caption{}
		\label{fig:bse5RUDAL_lineplot}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_RUDAL_lineplot}
		\caption{}
		\label{fig:ddccRUDAL_lineplot}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/distribution_parity}
		\caption{}
		\label{fig:BSE5_distribution_parity}
	\end{subfigure}
	\hfill

	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/distribution_parity}
		\caption{}
		\label{fig:ddccdistribution_parity}
	\end{subfigure}	
	\caption{Line plots to highlight the performance as the number of ansatz layers (AL; left) and re-upload depth (RUD; right) increases for the 5 qubits (a) BSE49 and (b) DDCC data. Additionally, regression parity plots for each AL and RUD are shown for the (c) BSE49 and (d) DDCC data, where the x-axis denotes the predicted target values and the y-axis denotes the reference data.}
	\label{fig:BSE_distribution_parity}	
\end{figure}

@@

Following the evaluation of the learning curves, we decided to only run the A2{\_}HWE-CNOT circuit on real quantum hardware.
Before running on the real hardware (\textit{ibm{\_}quebec}), we optimize the number of shots, the resilience level, and optimization using the \textit{FakeQuebec} backend with the \textit{qiskit-aer} plugin.
Based on the results from the observations provided by the state vector learning curves (Fig. \ref{fig:ddcclearningcurves}), along with considerations regarding the computational cost associated with the query time to the IBM Quantum platform, number of training points ($N_{\text{occ}}^{2}N_{\text{virt}}^{2}$ $t_{2}$-amplitudes), along with the potential number of device executions associated with the number of shots, resilience (error mitigation) and optimization levels the total number of training data is reduced from 80\% to 10\% of the total training data.
We validated this assumption using the \textit{ibm{\_}quebec} device using Twirled Readout Error eXtinction (TREX) \cite{van_den_berg_model-free_2022} error mitigation (resilience level 1), an optimization level of 2, with 3072 shots for both 10\%  and 80\% of the training data.
Using one iteration, we found that one iteration using 10\% of the training data took 1824 s of wall time, versus one iteration using 80\% of the training data taking 25004 s of wall time.
For a model that requires 500 or 1000 iterations to train, this would require approximately 145 to 289 days to train an end-to-end model using \textit{ibm{\_}quebec} using 80\% of the training data, versus 10 to 21 days required to train the model containing 10\% of the total data.
While this is still much longer than the classical models, the value of running on real hardware is still a worthwhile experiment to show how these models behave on the current generation of hardware.

To efficiently run the DDCC model on \textit{ibm{\_}quebec} the data is split into batches of 4 molecules per sample with 64 $t_{2}$-amplitudes each.
The motivation for batching the data is to reduce the number of circuit executions per primitive unified blocs (PUBs).
We found that for 500 iterations, an optimization level of X, with a resilience level of Y, using Z shots was adequate to reach satisfactory model performance using the \textit{FakeQuebec} backend (\textbf{Update these figures} Figs. \ref{fig:optresfakequebec} and \ref{fig:shotsfakequebec}).
When applied on \textit{ibm{\_}quebec} we found that the models had R$^{2}$s of X and Y, for the training and test set, respectively, and took a total of X QPU time and Y wall time to complete.

@@

level 1 readout error mitigation and measurement twirling using Twirled Readout Error eXtinction (TREX) \cite{van_den_berg_model-free_2022}
level 2 level 1 + gate twirling and zero noise extrapolation (ZNE)\cite{kandala_error_2019,li_efficient_2017,temme_error_2017}
